---
title: "BRFSS data exploration"
author: "Maria Guzman"
date: "11/18/2020"
output: github_document
---

### About BRFSS

*"The Behavioral Risk Factor Surveillance System (BRFSS) is the nation’s premier system of health-related telephone surveys that collect state data about U.S. residents regarding their health-related risk behaviors, chronic health conditions, and use of preventive services. Established in 1984 with 15 states, BRFSS now collects data in all 50 states as well as the District of Columbia and three U.S. territories. BRFSS completes more than 400,000 adult interviews each year, making it the largest continuously conducted health survey system in the world."*

```{r setup, include=FALSE}
library(tidyverse)
library(haven) #since this data is in xpt format
library(plotly)
library(viridis)

```

First, we need to import the data using read_xpt

```{r data import}

brfss_df =
  read_xpt("./final_project_large_data/LLCP2019.xpt") %>%
  janitor::clean_names()%>%
  mutate(
    state = as.factor(state),
    flushot_yn = as.factor(flushot7)
  )%>%
  mutate(
    state = recode(state,
         "1" = "Alabama",
         "2" = "Alaska",
         "4" = "Arizona",
         "5" = "Arkansas",
         "6" = "California",
         "8" = "Colorado",
         "9" = "Connecticut",
         "10" = "Delaware",
         "11" = "District of Columbia",
         "12" = "Florida",
         "13" = "Georgia",
         "15" = "Hawaii",
         "16" = "Idaho",
         "17" = "Illinois",
         "18" = "Indiana",
         "19" = "Iowa",
         "20" = "Kansas",
         "21" = "Kentucky",
         "22" = "Louisiana",
         "23" = "Maine",
         "24" = "Maryland",
         "25" = "Massachusetts",
         "26" = "Michigan",
         "27" = "Minnesota",
         "28" = "Mississippi",
         "29" = "Missouri",
         "30" = "Montana",
         "31" = "Nebraska",
         "32" = "Nevada",
         "33" = "New Hampshire",
         "35" = "New Mexico",
         "36" = "New York",
         "37" = "North Carolina",
         "38" = "North Dakota",
         "39" = "Ohio",
         "40" = "Oklahoma",
         "41" = "Oregon",
         "42" = "Pennsylvania",
         "44" = "Rhode Island",
         "45" = "South Carolina",
         "46" = "South Dakota",
         "47" = "Tennessee",
         "48" = "Texas",
         "49" = "Utah",
         "50" = "Vermont",
         "51" = "Virginia",
         "53" = "Washington",
         "54" = "West Virginia",
         "55" = "Wisconsin",
         "56" = "Wyoming",
         "66" = "Guam",
         "72" = "Puerto Rico"),
    flushot_yn = recode(flushot_yn,
          "1" = "Yes",
          "2" = "No",
          "7" = "Don't know/not sure",
          "9" = "Refused")
  )


#head(brfss_df)

```
per the environment and the code above, we can see that there are 418,268 observations and 342 variables. Upon first glance,
the names do not seem that intuitive, so we will have to refer to the codebook at the following link: https://www.cdc.gov/brfss/annual_data/2019/pdf/codebook19_llcp-v2-508.HTML

Based, off of this codebook and our interest in vaccines, I will peruse the data and select:
-a participant id  
-NATIONAL  
-demographic variables
-variables pertaining to flu vaccine


#### poking through the codebook and the clean names

Yesterday, I went through the codebook online and pulled the variables of interest. The purpose of creating this codebook,
which is in the project file called "BRFSS_codebook_streamlined", is to (1) compare variables with Ronae's national survey,
which seems to have a lot of overlap with this survey but may be more comprehensive. This comparison will lead us to (2) deciding
which dataset to keep and how to split up the larger dataset, if necessary. Finally, before our meeting today I will (3) perform
some preliminary data exploration and visualization now that I know which variables to select.

```{r data_streamline}

 
brfss_df %>%
  select(seqno, state, sexvar, genhlth, physhlth, menthlth, hlthpln1, persdoc2, medcost, checkup1,
         flushot7, imfvpla1, urbstat) %>%
  mutate(
    flu_vacc_loc = as.factor(imfvpla1)
  ) %>%
  drop_na(flu_vacc_loc) %>%
  mutate(
    flu_vacc_loc = recode(flu_vacc_loc,
                          '1' = "Doctor's office/HMO",
                          '2' = "Health dept",
                          '3' = "Clinic/health center",
                          '4' = "Senior, rec, or community center",
                          '5' = "Supermarket/pharmacy",
                          '6' = "Hospital",
                          '7' = "ER",
                          '8' = "Workplace",
                          '9' = "Elsewhere",
                          '10' = "Canada/Mexico",
                          '11' = "A school",
                          '77' = "Don't know/not sure",
                          '99' = "Refused"),
  )%>% 
  count(flu_vacc_loc)%>%
  mutate(flu_vacc_loc = fct_reorder(flu_vacc_loc, n)) %>%
    plot_ly(x = ~flu_vacc_loc, y = ~n, color = ~flu_vacc_loc, type = "bar", colors = "plasma") %>%
  layout(
    xaxis = list(title = "Flu vaccine location"),
    yaxis = list(title = "Count (n)")
  )
  

```
plot: flushot yes no by insurance yes no

```{r}

flushot_insurance =
brfss_df %>%
  mutate(
    insurance = as.factor(hlthpln1)
  ) %>%
  drop_na(insurance, flushot_yn) %>%
  mutate(
    insurance = recode(insurance,
                          '1' = 'Yes',
                          '2' = 'No')
  )%>% 
  filter(flushot_yn != "Don't know/not sure")%>%
  filter(flushot_yn != "Refused")%>%
  filter(insurance != "7")%>%
  filter(insurance != "9")%>%
  count(insurance, flushot_yn)%>%
  group_by(insurance, flushot_yn) %>%
  ggplot(aes(x = insurance, y = n, fill = flushot_yn))+
  geom_bar(position = "stack", stat = "identity")
  
ggplotly(flushot_insurance)

```

Ok finally here's a little plot of the counts of where people got the vaccine. How many people actually responded yes?

```{r}
library(wesanderson)

#Midwest
plot_midwest =
  brfss_df %>%
  filter(
    state == c("Illinois", "Indiana", "Iowa", "Kansas", "Michigan", "Minnesota",
               "Missouri", "Nebraska", "North Dakota", "Ohio", "South Dakota",
               "Wisconsin")
    )%>%
  drop_na(flushot_yn, state) %>%
  select (state, flushot_yn) %>%
  group_by(state, flushot_yn)%>%
  summarize(n_obs = n()) %>%
  mutate(
    pct = n_obs/sum(n_obs)
  )%>%
  filter(flushot_yn != "Don't know/not sure")%>%
  filter(flushot_yn != "refused")%>%
  ggplot(aes(fill = flushot_yn, y = pct, x = state))+
  geom_bar(position="dodge", stat = "identity")+
  ylim(0, 1.0)+
  theme_bw()+
  labs(
    x = "State",
    y = "% Response",
    title = "Frequency of flu vaccinations in the Midwestern region of the U.S"
  )+
  scale_fill_discrete(name = "Response")+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1))
 
ggplotly(plot_midwest) 
 

```


```{r}

plot_northeast =
 brfss_df %>%
  filter(
    state == c("Connecticut","District of Columbia", "Delaware","Maine", "Massachusetts", "Maryland",
               "New Hampshire", "New Jersey", "New York", "Pennsylvania",
               "Rhode Island", "Vermont")
    )%>%
  drop_na(flushot_yn, state) %>%
  select (state, flushot_yn) %>%
  group_by(state, flushot_yn)%>%
  summarize(n_obs = n()) %>%
  mutate(
    pct = n_obs/sum(n_obs)
  )%>%
  filter(flushot_yn != "Don't know/not sure")%>%
  filter(flushot_yn != "refused")%>%
  ggplot(aes(fill = flushot_yn, y = pct, x = state))+
  geom_bar(position="dodge", stat = "identity")+
  ylim(0, 1.0)+
  theme_bw()+
  labs(
    x = "State",
    y = "% Response",
    title = "Frequency of flu vaccinations in the Northeastern region of the U.S."
  )+
  scale_fill_discrete(name = "Response")+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1)) 

ggplotly(plot_northeast)

  
```


```{r}
plot_west =
 brfss_df %>%
  filter(
    state == c("Alaska", "Arizona", "California", "Colorado","Hawaii", "Idaho", "Montana",
               "Nevada", "New Mexico", "Oregon", "Utah", "Washington",
               "Wyoming")
    )%>%
  drop_na(flushot_yn, state) %>%
  select (state, flushot_yn) %>%
  group_by(state, flushot_yn)%>%
  summarize(n_obs = n()) %>%
  mutate(
    pct = n_obs/sum(n_obs)
  )%>%
  filter(flushot_yn != "Don't know/not sure")%>%
  filter(flushot_yn != "refused")%>%
  ggplot(aes(fill = flushot_yn, y = pct, x = state))+
  geom_bar(position="dodge", stat = "identity")+
  ylim(0, 1.0)+
  theme_bw()+
  labs(
    x = "State",
    y = "% Response",
    title = "Frequency of flu vaccinations in the Western region of the U.S."
  )+
  scale_fill_discrete(name = "Response")+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1))

ggplotly(plot_west)

```

```{r}
plot_south =
 brfss_df %>%
  filter(
    state == c("Alabama","Arkansas", "Florida", "Kentucky", "Louisiana", 
               "Mississippi", "North Carolina", "Oklahoma", "South Carolina",
               "Tennessee", "Texas", "Virginia", "West Virginia")
    )%>%
  drop_na(flushot_yn, state) %>%
  select (state, flushot_yn) %>%
  group_by(state, flushot_yn)%>%
  summarize(n_obs = n()) %>%
  mutate(
    pct = n_obs/sum(n_obs)
  )%>%
  filter(flushot_yn != "Don't know/not sure")%>%
  filter(flushot_yn != "refused")%>%
  ggplot(aes(fill = flushot_yn, y = pct, x = state))+
  geom_bar(position="dodge", stat = "identity")+
  ylim(0, 1.0)+
  theme_bw()+
  labs(
    x = "State",
    y = "% Response",
    title = "Frequency of flu vaccinations in the Southern region of the U.S."
  )+
  scale_fill_discrete(name = "Response")+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1))

ggplotly(plot_south)

```

#### Will try to iterate over datasets from 2015 to now and see if I can plot a trend
hopefully the code will be similar to what we did in the last homework

```{r}
goal_trend = 
list.files(path = "./final_project_large_data", full.names=FALSE)%>%
  tibble::enframe(name = NULL)

```

Data too large :-(
```{r}
# goal_trend_df =
#   goal_trend %>%
#   mutate(
#     data = map(value, ~read_xpt(str_c("./final_project_large_data/", .x)))
#   )%>%
#   unnest(cols = data)%>%
#   separate(value, into=c("survey", "year"))

```

#### I want to try and scrape data from twitter HA!

```{r}
library(rvest)
library(httr)
library(rtweet)
library(httpuv)
#look at twitr

url = "https://twitter.com/search?q=flu%20vaccines&src=typeahead_click"
twitter_html = read_html(url)%>%
  html_nodes(css = "table")

tweets = search_tweets(q = "#Vaccine", 
                        n = 18000,
                        include_rts = FALSE,
                        `-filter` = "replies",
                        lang = "en")

ts_plot(tweets, "hours") +
  labs(x = NULL, y = NULL,
       title = "Frequency of tweets with a #Vaccine hashtag",
       subtitle = paste0(format(min(tweets$created_at), "%d %B %Y"), " to ", format(max(tweets$created_at),"%d %B %Y")),
       caption = "Data collected from Twitter's REST API via rtweet") +
  theme_minimal()

```
