---
title: "National Health Survey"
author: "Ronae McLin"
date: "11/16/2020"
output: github_document
---

## importing and tidying the National Health Survey data 

***focus is on flu vaccines***

```{r}
library(tidyverse)
library(plotly)
library(wesanderson)
library(gganimate)
library(gifski)
```

not really sure what this data set contains. need to look into this more

adult sample for 2019

** there is ALOT of information and variables within this dataset.  we can take this into many different directions **
```{r}
  adult_df = 
  read_csv("adult19.csv")
```


- was SA pregnant last flu season?
- flu vaccine
- month of last flu vaccine,
- year of last flu vaccine
- was flu shot before or during pregnancy?
- earlier pregnancy and flu vaccine
- etc....

there is an interesting category pertaining to mobility, this could be helpful if we end up discussing accessbility to loaction to obtain a flu shot

there is also interesting category of variables pertaining to insurance plans and coverage.

interesting category of variables pertaining to utilization of healthcare services, outside of the scope of only flu vaccines. 
```{r}
flu_adult_df =
  adult_df %>% 
  janitor::clean_names() %>% 
  select(
    srvy_yr,agep_a, age65, sex_a, educ_a,hisp_a, hispallp_a, hisdetp_a,phstat_a,bmicat_a, notcov_a, cover_a, usualpl_a, usplkind_a,  pregfluyr_a, shtflu12m_a, shtflum_a, shtfluy_a, flupreg2_a, raceallp_a, pregnow_a, disab3_a, socerrnds_a, cover65_a, lastdr_a, emerg12mtc_a, urgnt12mtc_a, natusborn_a,citznstp_a, povrattc_a, pregnow_a, paybll12m_a, wtfa_a
  ) %>% 
  mutate(
    weight = wtfa_a, 
    age =  as.factor(agep_a), 
      age65 = case_when(
      age65 == "1" ~ "Less than 65",
      age65 == "2" ~ "65 or older",
      age65 == "7" ~ "No response",
      age65 == "8" ~ "Not applicable",
      age65 == "9" ~ "Don't know"),
      
      sex = case_when(
      sex_a == "1" ~ "Male",
      sex_a == "2" ~ "Female",
      sex_a == "7" ~ "No response",
      sex_a == "8" ~ "Not applicable",
      sex_a == "9" ~ "Don't know"),
      
      education = case_when(
      educ_a == "0" ~ "never attended",
      educ_a == "1" ~ "Grades 1-11",
      educ_a == "2" ~ "12th grade, no dipolma",
      educ_a == "3" ~ "GED",
      educ_a == "4" ~ "HS graduate",
      educ_a == "5" ~ "College, no degree",
      educ_a == "6" ~ "Associate degree-tech",
      educ_a == "7" ~ "Associate degree-aca",
      educ_a == "8" ~ "Bachelor's degree",
      educ_a == "9" ~ "Master's degree",
      educ_a == "10" ~ "Professional school",
      educ_a == "11" ~ "Doctoral degree",
      educ_a == "97" ~ "No response",
      educ_a == "98" ~ "Not applicable",
      educ_a == "99" ~ "Don't know"),
      
      hispanic = case_when(
      hisp_a == "1" ~ "Yes",
      hisp_a == "2" ~ "No",
     hisp_a == "7" ~ "No response",
     hisp_a == "8" ~ "Not applicable",
      hisp_a == "9" ~ "Don't know"),
     
     nh_race = case_when(
      hispallp_a == "1" ~ "hispanic",
      hispallp_a == "2" ~ "NH white",
     hispallp_a == "3" ~ "NH black", 
     hispallp_a == "4" ~ "NH asian",
     hispallp_a == "5" ~ "NH AIAN",
      hispallp_a == "6" ~ "NH AIAN and other",
      hispallp_a == "7" ~ "multiple races",
       hispallp_a == "97" ~ "No response",
     hispallp_a == "98" ~ "Not applicable",
      hispallp_a == "99" ~ "Don't know"),
     
      hispanic_type = case_when(
      hisdetp_a == "1" ~ "hispanic mexican",
      hisdetp_a == "2" ~ "hispanic all others",
     hisdetp_a == "3" ~ "Not hispanic", 
        hisdetp_a == "7" ~ "No response",
     hisdetp_a == "8" ~ "Not applicable",
      hisdetp_a == "9" ~ "Don't know"),
     
     health =  case_when(
      phstat_a == "1" ~ "excellent",
      phstat_a == "2" ~ "very good",
      phstat_a == "3" ~ "good",
      phstat_a == "4" ~ "fair",
      phstat_a == "5" ~ "poor",
      phstat_a == "7" ~ "no response",
      phstat_a == "8" ~ "not applicable",
      phstat_a == "9" ~ "Don't know"),
     
      bmi =  case_when(
      bmicat_a == "1" ~ "underweight",
      bmicat_a == "2" ~ "healthy weight",
       bmicat_a == "3" ~ "overweight",
       bmicat_a == "4" ~ "obese",
       bmicat_a == "9" ~ "unknown"),
     
  insurance = case_when(
    notcov_a == "1" ~ "not covered",
    notcov_a == "2" ~ "covered",
    notcov_a == "7" ~ "no response",
    notcov_a == "8" ~ "not applicable",
    notcov_a == "9" ~ "don't know"),
  
    coverage = case_when(
      cover_a == "1" ~ "private",
      cover_a == "2" ~ "medicaid/other public",
      cover_a == "3" ~ "other",
      cover_a == "4" ~ "uninsured",
      cover_a == "5" ~ "don't know"),
  
  facility_type = case_when(
    usplkind_a == "1" ~ "Doctor office/health center",
    usplkind_a == "2" ~ "Urgent care/clinic in store",
    usplkind_a == "3" ~ "emergency room",
    usplkind_a == "4" ~ "VA center ",
    usplkind_a == "5" ~ "other",
    usplkind_a == "6" ~ "varies",
    usplkind_a == "8" ~ "not applicable",
    usplkind_a == "9" ~ "don't know"),
  
  pregfluyr_a = case_when(
  pregfluyr_a == "1" ~ "yes",
  pregfluyr_a == "2" ~ "no",
  pregfluyr_a == "7" ~ "no response",
  pregfluyr_a == "8" ~ "not applicable",
  pregfluyr_a == "9" ~ "don't know"),
  
  flu_shot  = case_when(
  shtflu12m_a == "1" ~ "yes",
  shtflu12m_a == "2" ~ "no",
  shtflu12m_a == "7" ~ "no response",
  shtflu12m_a == "8" ~ "not applicable",
  shtflu12m_a == "9" ~ "don't know"),
  
    flu_shot_month = case_when(
      shtflum_a == "1" ~ "January ",
      shtflum_a == "2" ~ "February",
      shtflum_a == "3" ~ "March",
      shtflum_a == "4" ~ "April",
      shtflum_a == "5" ~ "May",
      shtflum_a == "6" ~ "June ",
      shtflum_a == "7" ~ "July",
      shtflum_a == "8" ~ "August",
      shtflum_a == "9" ~ "September",
      shtflum_a == "10" ~ "October",
      shtflum_a == "11" ~ "November",
      shtflum_a == "12" ~ "December",
      shtflum_a == "97" ~ "No response",
      shtflum_a == "98" ~ "Not applicable",
      shtflum_a == "99" ~ "Don't know"),
  
   flu_shot_year = case_when(
    shtfluy_a == "9997" ~ "no response",
    shtfluy_a == "9998" ~ "not applicable",
    shtfluy_a == "9999" ~ "don't know"),
  
  flushot_preg = case_when(
    flupreg2_a == "1" ~ "before pregnancy",
    flupreg2_a == "2" ~ "during pregnancy",
    flupreg2_a == "3" ~ "after pregnancy",
    flupreg2_a == "7" ~ "no response",
    flupreg2_a == "8" ~ "not applicable",
    flupreg2_a == "9" ~ "don't know"),

usual_place = case_when(
    usualpl_a == "1" ~ "yes",
    usualpl_a == "2" ~ "no",
    usualpl_a == "3" ~ "more than one place",
    usualpl_a == "7" ~ "no response",
    usualpl_a == "8" ~ "not applicable",
    usualpl_a == "9" ~ "don't know"))


```


    
```{r}
flu_adult_df = 
flu_adult_df %>% 
   mutate(
      pregnant = case_when(
      pregnow_a == "1" ~ "yes",
      pregnow_a == "2" ~ "no",
      pregnow_a == "7" ~ "No response",
      pregnow_a == "8" ~ "Not applicable",
      pregnow_a == "9" ~ "Don't know"),
      
      disability = case_when(
      disab3_a == "1" ~ "yes",
      disab3_a == "2" ~ "no",
      disab3_a == "9" ~ "don't know"),
      
      errands_alone_diff = case_when(
    socerrnds_a == "1" ~ "none",
    socerrnds_a == "2" ~ "some",
    socerrnds_a == "3" ~ "a lot",
    socerrnds_a == "4" ~ "cannot do at all",
    socerrnds_a == "7" ~ "no response",
    socerrnds_a == "8" ~ "not applicable",
    socerrnds_a == "9" ~ "don't know"),
    
    coverage_65 = case_when(
    cover65_a == "1" ~ "private",
    cover65_a == "2" ~ "dual eligible",
    cover65_a == "3" ~ "medicare advantage",
    cover65_a == "4" ~ "medicare only",
    cover65_a == "5" ~ "other",
    cover65_a == "6" ~ "uninsured",
    cover65_a == "7" ~ "don't know"))
      
      
      
      
```


```{r}
flu_adult_df = 
flu_adult_df %>% 
  mutate(
     last_drvisit = case_when(
    lastdr_a == "0" ~ "never",
    lastdr_a == "1" ~ "within past year",
    lastdr_a == "2" ~ "within last 2 years",
    lastdr_a == "3" ~ "within last 3 years",
    lastdr_a == "4" ~ "within last 5 years",
    lastdr_a == "5" ~ "within last 10 years",
    lastdr_a == "6" ~ "10+ years",
    lastdr_a == "7" ~ "refused",
    lastdr_a == "8" ~ "not applicable",
    lastdr_a == "9" ~ "don't know"),
    
    emergency_visits = case_when(
    emerg12mtc_a == "0" ~ "never",
    emerg12mtc_a == "1" ~ "once",
    emerg12mtc_a == "2" ~ "twice",
    emerg12mtc_a == "3" ~ "thrice",
    emerg12mtc_a == "4" ~ "4+ times",
    emerg12mtc_a == "7" ~ "refused",
    emerg12mtc_a == "8" ~ "not applicable",
    emerg12mtc_a == "9" ~ "don't know"),
    
     urgentcare_visits = case_when(
    urgnt12mtc_a == "0" ~ "never",
    urgnt12mtc_a == "1" ~ "once",
    urgnt12mtc_a == "2" ~ "twice",
    urgnt12mtc_a == "3" ~ "thrice",
    urgnt12mtc_a == "4" ~ "4 times",
    urgnt12mtc_a == "5" ~ "5+ times",
    urgnt12mtc_a == "7" ~ "refused",
    urgnt12mtc_a == "8" ~ "not applicable",
    urgnt12mtc_a == "9" ~ "don't know")
  )
```

 
```{r}
flu_adult_df = 
flu_adult_df %>% 
  mutate(
     usborn = case_when(
      natusborn_a == "1" ~ "yes",
      natusborn_a == "2" ~ "no",
      natusborn_a == "7" ~ "No response",
      natusborn_a == "8" ~ "Not applicable",
      natusborn_a == "9" ~ "Don't know"),
     
     probs_paybill = case_when(
      paybll12m_a == "1" ~ "yes",
      paybll12m_a == "2" ~ "no",
      paybll12m_a == "7" ~ "No response",
      paybll12m_a == "8" ~ "Not applicable",
      paybll12m_a == "9" ~ "Don't know"))
```

last tidy up
```{r}
flu_adult_df = 
flu_adult_df %>% 
  select(-agep_a, -sex_a, -educ_a,-hisp_a,-phstat_a,-bmicat_a, -cover_a, -usplkind_a,  -pregfluyr_a, -shtflu12m_a, -shtflum_a, -shtfluy_a, -flupreg2_a, -pregnow_a, -disab3_a, -socerrnds_a, -cover65_a, -lastdr_a, -emerg12mtc_a, -urgnt12mtc_a, -natusborn_a, -pregnow_a, -paybll12m_a
  )
```

```{r}
flu_adult_df = 
  flu_adult_df %>% 
  select(
    -hispallp_a, -hisdetp_a, -citznstp_a, -povrattc_a, -raceallp_a, -usualpl_a, -notcov_a, -wtfa_a
  )
```


```{r}
flu_adult_df = 
  flu_adult_df %>% 
  mutate(
    sex = as.factor(sex),
    nh_race = as.factor(nh_race),
    health = as.factor(health),
    bmi = as.factor(bmi)
  )
```

```{r}
flu_adult_df = 
  flu_adult_df %>% 
  mutate(
    coverage = as.factor(coverage),
    education = as.factor(education),
    insurance = as.factor(insurance),
    flu_shot = as.factor(flu_shot),
    flu_shot_month = as.factor(flu_shot_month))
```

colors
```{r}
buda <- wes_palette("GrandBudapest2", 21, type = "continuous")
```

plots bebe

# basic demographics

## Looking at biological sex

Starting off we can see that females are more likely to get a flu shot than not, this does not appear to be the same trend with men.  Men are more likely to not receive a flu shot than to get one. 
```{r}

flu_adult_df %>% 
  filter(flu_shot != "no response") %>% 
  filter(flu_shot != "not applicable") %>% 
  filter(flu_shot != "don't know") %>% 
  filter(sex != "No response") %>% 
  ggplot(aes( x = flu_shot, fill = sex)) + geom_bar() + facet_grid(.~ sex) + scale_fill_manual(values = wes_palette("GrandBudapest2"))
```


let's look at the months when shot are most likely to happen, and who are most likely to get shots during this time.

for those who have flu shot under 65. 

not excited by this plot
```{r}
flu_adult_df %>% 
 filter(flu_shot_month != "No response") %>% 
  filter(flu_shot_month != "Not applicable") %>% 
  filter(flu_shot_month != "Don't know") %>% 
  plot_ly(x = ~education, y = ~health, type = "bar", mode = "markers", color = ~coverage) 

```


## when are flu shots occurring? and insurance type

Let's look at insurance in relation to flu shots. Most of those receiving a flu shot are covered under private insurance.  Shots typically occur in October, with November close behind.  

**factor reorder by time of months**
```{r}
flu_adult_df %>% 
  filter(flu_shot_month != "No response") %>% 
  filter(flu_shot_month != "Not applicable") %>% 
  filter(flu_shot_month != "Don't know") %>% 
  filter(coverage != "don't know") %>%
  filter(coverage != "NA") %>% 
  ggplot(aes(x = coverage, y = flu_shot_month, color = coverage, alpha = .04)) + geom_jitter() + scale_fill_manual(values = wes_palette("GrandBudapest2"))
```

## money and bills

for those who didn't receive a flu shot, most respondents did not have difficulty paying their hospital bill.  

**factor reorder here babes**
```{r}
flu_adult_df %>% 
   filter(flu_shot == "no") %>%
    filter(probs_paybill != "Don't know") %>% 
  filter(probs_paybill != "No response") %>%
   filter(last_drvisit != "don't know") %>% 
  filter(last_drvisit != "refused") %>%
  ggplot(aes(fill = probs_paybill, y = last_drvisit)) + geom_bar() + scale_fill_manual(values = wes_palette("Royal2"))

```

## Race and shots

Age distribution for not getting a shot
```{r}
flu_adult_df %>% 
  filter(flu_shot == "no") %>% 
plot_ly(y = ~age, color = ~nh_race, type = "box", colors = "viridis") 
```

## Age distribution for getting a shot based on education

**factor reorder babes**

```{r}
flu_adult_df %>% 
  filter(flu_shot == "yes") %>% 
plot_ly(y = ~age, color = ~education, type = "box", colors = "viridis") 
```

Seems like people are generally older when they receive a flu vaccine. 


## plot for pregnant people and flu shots

```{r}

```


## plot for emergency care and urgent care visits

```{r}
flu_adult_df %>% 
  ggplot(aes(x = emergency_visits, y = age)) + geom_density() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```



## hispanic

when we look at covid rates, we notice that black and brown communities are disporitionaitely affected by infection.  Let's look at this for Hispanice groups in regards to flu and health

```{r}

```

## general health and BMI

```{r}

```


## disabilitiies babes


