---
title: "NYC Flu Vaccination Locations"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(httr)
library(rvest)
library(plotly)
library(leaflet)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Data Description

This data set contains information about the locations in NYC where individuals can get a flu shot. This data was collected by NYC DOHMH and is hosted through NYC OpenData. You can view the data set page [here](https://data.cityofnewyork.us/Health/New-York-City-Locations-Providing-Seasonal-Flu-Vac/w9ei-idxz). Each location where people can obtain an influenza vaccine is listed along with business name, location, and whether they accept insurance, walk-in appointments, or vaccinate children.

We will start by importing and tidying the data set.

```{r import_tidy, message = FALSE}
flu_vax = 
  GET("https://data.cityofnewyork.us/resource/w9ei-idxz.csv") %>% 
  content("parsed")

vax_tidy = 
  flu_vax %>% 
  select(objectid, facility_name, walk_in, insurance, children, borough, zip_code, latitude, longitude) %>% 
  mutate(borough = str_to_title(borough))
```

### Distribution of Vaccination Sites

Let's look at the spatial distribution of locations using `plotly`.

```{r plotly, warning = FALSE}
vax_plotly = 
  vax_tidy %>% 
  mutate(text_label = str_c(facility_name, "\nZip Code: ", zip_code)) %>% 
  plot_ly(
    x = ~longitude, y = ~latitude, color = ~borough,
    text = ~text_label, alpha = 0.7, type = "scatter", mode = "markers"
  ) 

vax_plotly
```

That's pretty helpful, but we can drop the data onto a real map using `leaflet`.

```{r leaflet}
vax_leaf = 
  vax_tidy %>% 
  mutate(text_label = str_c(facility_name, ", ", zip_code)) %>%
  leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addCircleMarkers(lat = ~latitude, lng = ~longitude, radius = 0.1, 
                   label = ~text_label)

vax_leaf
```

We can see that there is a higher concentration of vaccination sites in Manhattan, which we would expect due to population density. Of all locations, `r format(round((nrow(filter(vax_tidy, walk_in == "Yes")) / nrow(vax_tidy)) * 100, 2))`% accept walk-in visits, `r format(round((nrow(filter(vax_tidy, insurance == "Yes")) / nrow(vax_tidy)) * 100, 2))`% accept insurance, and `r format(round((nrow(filter(vax_tidy, children == "Yes")) / nrow(vax_tidy)) * 100, 2))`% vaccinate children. Unfortunately, the data dictionary on NYC OpenData is inaccurate, so we do not know whether `children` means that flu vaccinations are administered to children at a location, or if they also administer other childhood vaccinations (like TDAP, MMR, etc). 
